---
title: "Homework #1: Jump Start with RMarkdown"
subtitle: "Monte Carlos Approximation"
author:
  - Yuqian Liu^[<yuqian.liu@uconn.edu>; Ph.D. student at
    Department of Mechanical Engineering, University of Connecticut.]
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: template.bib
biblio-style: datalab
keywords: Approximation, Monte Carlo, Normal Destribution, R Markdown, bookdown
# keywords set in YAML header here only go to the properties of the PDF output
# the keywords that appear in PDF output are set in latex/before_body.tex
output:
# bookdown::pdf_document2
  bookdown::html_document2
abstract: |
    This is the first homework for STAT5361 Statistical Computat     ion which is 
    designed for getting familiar with RMarkdown. In this homework, I will 
    approximate one integration by Monte Carlo method with sampling 
    from Normal distribution function N(0,1). The experiments would be 
    conducted at 3 different sample sizes, with 9 different upper limits. 
    The key-step derivation and results will be presented. Ture values will
    also be calculated as reference.
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")
source("main_Yuqian.R")

## specify the packages needed
pkgs <- c("splines2", "DT", "webshot", "leaflet")
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## get output format in case something needs extra effort
outFormat <- knitr::opts_knit$get("rmarkdown.pandoc.to")
## "latex" or "html"

## for latex and html output
isHtml <- identical(outFormat, "html")
isLatex <- identical(outFormat, "latex")
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```


# Introduction {#sec:intro}

Homework 1 is using the template provided by Wenjie Wang to organize ...
This document is designed as a template for data science lab projects.  However,
it can also be used as a general template in `r proglang("R Markdown")` for a
single document.


Introduce Monte Carlo method


The rest of this homework is organized as follows: Section \@ref(sec:problem)
will give problem discription. In Section \@ref(sec:math),
mathematical derivation will be illustrated. The procedure of solving this
integration will be presented step by step. The results of different scenarios
will be given in Section \@ref(sec:results). The ture value of the integrals 
will also be calculated as references. After repeating each case 100 times, 
Section \@ref(sec:figure) will give box plots of the bias. The code chunk of this
homework's calculation is given in Section \@ref(sec:code). In the end, summary and
discussion will be presented in Section \@ref(sec:summary).



# Problem Formulation{#sec:problem}


Consider approximation of the distribution function of $N(0,1)$,
\begin{align}
    \Phi (t) = \int_\infty ^t {\frac{1}{\sqrt{2\pi }}} {e^{ - {y^2}/2}}dy
    (\#eq:integral)
\end{align}
by the Monte Carlo methods:
\begin{align}
   \hat \Phi (t) = \frac{1}{n}\sum\limits_{i = 1}^n {I({X_i} \le t)}
   (\#eq:final)
\end{align}
where $X_i$'s are independent, identically distributed (i.i.d.) $N(0,1)$ variables. 


Experiment with the approximation at $n\in\{{10^2},{10^3},{10^4}\}$ at 
$t\in\{0.0,0.67,0.84,1.28,1.65,2.32,2.58,3.09,3.72\}$ to form a table. The 
table should include the true value for comparison. Further repeat the experiment
100 times. Draw box plots of the bias at all $t$.



# Math Derivation {#sec:math}


Consider Equation \@ref(eq:integral) in this form,
\begin{align}
    \Phi (t) = \int_{ - \infty }^\infty  {h(y)g(y)} dy
\end{align}


Based on Monte Carlo method,


**Step 1**: Define $h(y) = \left\{ {\begin{array}{*{20}{c}} 1&{y \le t}\\ 0&{y > t} \end{array}} \right.$


**Step 2**: Define $g(y) = \frac{1}{{\sqrt {2\pi } }}{e^{ - {y^2}/2}}$.
Thus, the probability distribution function is,
\begin{align}
    p(y) = \frac{{g(y)}}{{\int_{ - \infty }^\infty  {g(y)dy} }} = \frac{1}{{\sqrt {2\pi } }}{e^{ - {y^2}/2}}
\end{align}


**Step 3**: Draw $n$ i.i.d. random samples from $p(y)$.


**Step 4**: Evaluate
\begin{align}
\hat \Phi (t) = C{\rm E}[h(x)] \approx \frac{C}{n}\sum\limits_{i = 1}^n {h({x_i})}  = \frac{1}{n}\sum\limits_{i = 1}^n {I({X_i} \le t)}
\end{align}
based on samples from **Step 3**.



# Results  {#sec:results}

The approximations with one-time run at $n\in\{{10^2},{10^3},{10^4}\}$ and 
$t\in\{0.0,0.67,0.84,1.28,1.65,2.32,2.58,3.09,3.72\}$ are  shown in Table
\@ref(tab:data1). The true values of integration are also provided for comparison.


<!-- source extracted from main_yuqian.R  -->

```{r data1} 
knitr::kable(final_data_1, caption = 'Ture Values and Prediction Values', booktabs = TRUE)
```



# Figures {#sec:figure}

After repeating each case 100 times, box plots are used to demonstrate results. Figure \@ref(fig:s100) 
shows predictions with 100 samples. Predictions with 1000 and 10000 samples are shown in Figure \@ref(fig:s1000)
and Figure \@ref(fig:s10000) respectively. One of boxplot code chunk is shown as follow.

<!-- caption of figure may be defined as follows: -->

(ref:s100) 100-run Predictions with 100 samples.

```{r s100, echo = TRUE, fig.cap = "(ref:s100)", fig.width = 8}
boxplot(Pred_100 ~ t, data = final_data_100, xlab = "t",ylab = "Predictions")
```
```{r s1000, echo = FALSE, fig.cap = "100-run Predictions with 1000 samples", fig.width = 8}
boxplot(Pred_1000 ~ t, data = final_data_100, xlab = "t",ylab = "Predictions")
```
```{r s10000, echo = FALSE, fig.cap = "100-run Predictions with 10000 samples", fig.width = 8}
boxplot(Pred_10000 ~ t, data = final_data_100, xlab = "t",ylab = "Predictions")
```

If we would like to see box plots of bias for case with different sample sizes, Figure \@ref(fig:bias100), 
\@ref(fig:bias1000) and \@ref(fig:bias10000) are given as follow.
```{r bias100, echo = FALSE, fig.cap = "Bias of 100-run Predictions with 100 samples", fig.width = 8}
boxplot(bias_100 ~ t, data = final_data_bias, xlab = "t",ylab = "Prediction Bias")
```
```{r bias1000, echo = FALSE, fig.cap = "Bias of 100-run Predictions with 1000 samples", fig.width = 8}
boxplot(bias_1000 ~ t, data = final_data_bias, xlab = "t",ylab = "Prediction Bias")
```
```{r bias10000, echo = FALSE, fig.cap = "Bias of 100-run Predictions with 10000 samples", fig.width = 8}
boxplot(bias_10000 ~ t, data = final_data_bias, xlab = "t",ylab = "Prediction Bias")
```


(ref:cap-miSplines) Quadratic M-spline (left) and I-spline (right) Bases with
three internal knots.



In the code chunk shown above, the chunk option `out.width = '45%'` and
`fig.show = 'hold'` were set so that the plots were placed side by side.  We may
set the chunk option `echo = FALSE` so that the code chunk generating the plots
are excluded from the output. Also, the chunk option `cache` can be set to be
`TRUE` for time-consuming code chunks once the code chunk is unlikely to be
modified.



# Code Chunk {#sec:code}

In addition to `r proglang("R")`, the code chunk can be written in a variety of
other languages, such as `r proglang("Bash")`, `r proglang("Python")`,
`r proglang("SAS")`, etc., by specifying the chunk option `engine`.
The following code chunk is one toy example written in `r proglang("Python 3")`.

<!-- note that we actually set eval to be FALSE here so that readers are able to
reproduce this template without Python 3. -->

```{r, eval = FALSE}
t <- c(0,0.67,0.84,1.28,1.65,2.32,2.58,3.09,3.72)
n <- c(100,1000,10000)

# true value
ture_Phi <- pnorm(t)

# Prediction function
Pred_Phi_100 <- function(x){
  y <- rnorm(100);
  m <- sum(y<=x);
  m/100
}

Pred_Phi_1000 <- function(x){
  y <- rnorm(1000);
  m <- sum(y<=x);
  m/1000
}

Pred_Phi_10000 <- function(x){
  y <- rnorm(10000);
  m <- sum(y<=x);
  m/10000
}

## One-time experiment
# try sapply function
Pred_1_100 <- sapply(t,Pred_Phi_100)
Pred_1_1000 <- sapply(t,Pred_Phi_1000)
Pred_1_10000 <- sapply(t,Pred_Phi_10000)
final_data_1 <- data.frame(t = t, `Ture Value`= ture_Phi, `n = 100` = Pred_1_100, `n = 1000` = Pred_1_1000,
                           `n = 10000` = Pred_1_10000, stringsAsFactors = FALSE, check.names = FALSE)

## 100-time experiment
# T <- matrix(rep(t,each = 100), nrow = 100)
T <- rep(t,each = 100)
# Ture_Phi <- matrix(rep(ture_Phi,each = 100), nrow = 100)
# n = 100
Pred_100_100 <- sapply(T,Pred_Phi_100)
Pred100_100 <- array(Pred_100_100,c(100,9)) #reshape

# n = 1000
Pred_100_1000 <- sapply(T,Pred_Phi_1000)
Pred100_1000 <- array(Pred_100_1000,c(100,9)) #reshape

# n = 10000
Pred_100_10000 <- sapply(T,Pred_Phi_10000)
Pred100_10000 <- array(Pred_100_10000,c(100,9)) #reshape

final_data_100 <- data.frame(t = T, Pred_100 = Pred_100_100, Pred_1000 = Pred_100_1000, Pred_10000 = Pred_100_10000)

## get predictions' bias
# mean of predictions
mean100 <- apply(Pred100_100,2,mean)
mean1000 <- apply(Pred100_1000,2,mean)
mean10000 <- apply(Pred100_10000,2,mean)
Mean100 <- rep(mean100,each = 100)
Mean1000 <- rep(mean1000,each = 100)
Mean10000 <- rep(mean10000,each = 100)
# bias
bias100 <- Pred_100_100 - Mean100
bias1000 <- Pred_100_1000 - Mean1000
bias10000 <- Pred_100_10000 - Mean10000
final_data_bias <- data.frame(t = T, bias_100 = bias100, bias_1000 = bias1000, bias_10000 = bias10000)
```


The chunk option `eval = FALSE` is set, in order to present the code without evaluation.



# Summary {#sec:summary}

In this homework, I am getting farmiliar with R lan
In summary, we provided this project template and reviewed most common
components and their syntax of writing a single `r proglang("R Markdown")`
document with the power and love of `r pkg("bookdown")` and many other fantastic
packages.


@xie2017bookdown provided a thorough introduction to `r pkg("bookdown")`
including more advanced customization and other output formats.
Additionally, the [manual][pandocManual] of `r pkg("Pandoc")` gives all the
available options that can be specified through the `r proglang("YAML")`
metadata section.


The template source and other associated files, such as BibTeX and CSS file, are
available at our GitHub repository
*dslab-templates*: <https://github.com/statds/dslab-templates>.


# Acknowledgment {-}

Thanks to Wenjie Wang's template and all the other authors and contributors.


# Reference {-}


[pandoc]: http://pandoc.org/
[pandocManual]: http://pandoc.org/MANUAL.html
[repo]: https://github.com/wenjie2wang/datalab-templates
[taskView]: https://cran.r-project.org/web/views/ReproducibleResearch.html
[shiny.io]: https://www.shinyapps.io/
[wenjie-stat.shinyapps]: https://wwenjie-stat.shinyapps.io/minisplines2
